### 2019-10-15 tesis_hiram/Acknowledgement/acknowledgement.tex
algo

% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}

And I would like to acknowledge Dr. Jos\'e Feliciano Ben\'itez Rubio for his contributions and advices in order to develop  and write this thesis. Also thank Dr. Dr. Javier Murillo and  Dr. Cristina Oropeza for their observations and advices in my formation. 
DR. Susana Alv\'arez for support during the process of the thesis. Acknowledgments to CONACYT for support this thesis with financial aid.

\end{acknowledgements}