### 2019-10-15 tesis_hiram/Abstract/abstract.tex
fixes

% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
\begin{abstract}
I present a study on the production of a Higgs boson plus a single top quark ($tH$) in same sign dimuon channel $\mu^\pm \mu^\pm$ using data published by the CMS experiment at CERN. By studying this
process I explore a Higgs production mechanism which has not yet been observed
experimentally. The expected sensitivity is estimated by using an Asimov dataset for 35.9 fb$^{-1}$ of proton-proton collisions and is extrapolated to the LHC high luminosity phase.
The sensivity is also studied for the model with inverted Yukawa coupling $k_t$=-1, where $k_t$ is the top-Higgs coupling parameter, for comparison with the Standard Model.
\end{abstract}