### 2019-11-24 tesis_hiram/Acknowledgement/acknowledgement.tex

% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}      
And I would like to acknowledge Dr. Jos\'e Feliciano Ben\'itez Rubio for his contributions and advises in order to develop  and write this thesis. Also thank Dr. Javier Murillo and  Dr. Cristina Oropeza for their observations and advises in my formation. 
Dr. Susana Alv\'arez for supporting during the process of the thesis. Acknowledgments to CONACYT for supporting this thesis with financial aid. 
\end{acknowledgements}