### 2019-11-24 tesis_hiram/Chapter1/chapter1.tex

	%!TEX root =../thesis.tex
%*******************************************************************************
%*********************************** First Chapter *****************************
%*******************************************************************************
%\renewcommand{\baselinestretch}{1.5}

\begin{chapter}{Introduction}

\section{General objective and motivations}

Particle physics is a field of physics that studies the particles that create matter and their interactions to very small scale, even smaller than atoms. 
Particle physics studies the fundamental building blocks of matter (fermions) and their interactions via force carriers (bosons). 
%The particles analyzed in particle physics are the fundamental particles that compose matter and the responsible for all forces of nature.
All those particles and their properties are grouped in a theory called the Standard Model (SM). SM is a theory that describes the fundamental forces (except gravity) and classifies the elementary particles in bosons (integer spin) and fermions (half spin).
The SM, created almost 50 years ago, has been the most successful theory in explaining the phenomena in the subatomic world.

One of the particles of the SM, the Higgs Boson, was recently discovered in 2012 in the ATLAS and CMS experiments at CERN. The discovery of Higgs boson in proton proton (pp) collisions marked a new era in particle physics due to its characteristic of giving matter the property of mass, according to previous theories and predictions. Since then, people around the world are working in the Large Hadron Collider (LHC) at CERN in order to detect the Higgs bosons in higher energies and different Higgs production processes. 
Also the Higgs boson is also important in cosmology, where theoretical investigations relate the Higgs boson to the cosmic inflation.

In recent years, there have been many studies of Higgs production processes at CERN in pp collisions, such as gluon fusion (ggF), VBF, WH, ZH and $t\bar{t}H$. 
In this work, we will analyze the production of a Higgs boson in association with a
single top quark (tH) in pp collisions with the CMS experiment of the LHC.

The motivation of the work is that this production mechanism of the Higgs boson has not been observed before by any experiment. The exploration of Higgs production on the tH channel is a relatively new subject. %The SM is used to study the tH channel and compare it with other Higgs production mechanisms results.
Many channels are yet out of
reach experimentally. At the same time, many theoretical
proposals are yet to be put to test. It is possible that studies on the Higgs boson give evidence of small deviations from the SM and test new physics that are beyond SM such as String Theory and Super-symmetry.

%********************************** % Third Section *************************************
\section{The Standard Model} %Section - 1.3 
The SM, created in the decade of the seventies, is a successful theory that explains the structure of matter and the interactions of three of the four fundamental forces which are electromagnetic, strong and weak. The gravitational force is yet to be included due to the scale of the interaction with respect to the other 3. All matter is composed of three kinds of particles: leptons, quarks and bosons, also called mediators.

For the leptons we have electrons ($e$), muons ($\mu$), and taus ($\tau$) with charge -1 and their neutrinos $\nu_e, \nu_\mu, \nu_\tau$, are chargeless particles. %Leptons have their respective antiparticle; same structure with opposite charge. 
 All of them have a spin (intrinsic angular momentum) of $\frac{1}{2}$\cite{griff}. % In case of neutrinos, so far experimentalists have found these have very small masses, but in the theoretical scheme, they are massless. \\
 Quarks combine to form composite particles called hadrons. Quarks cannot be found in isolation, they always exist in form of hadrons that can be mesons or baryons. Mesons are made up of a quark-antiquark pair, such as pion ($\pi$) and kaon ($K$). Mesons were theorized by Yukawa in 1937 and discovered in 1947 by Powell\cite{griff}.
Baryons are made of 3 quarks. Examples of baryons are protons and neutrons. Both types of particles can interact via the strong force. Since quarks and gluons only exist in bound, it is impossible to observe isolated quarks except the top quark, which decays before it has time to form a
bound state \cite{gross}.

Quarks were first theorized by Murray Gell-Mann and 
George Zweig in 1964. They were introduced as parts of an ordering scheme for hadrons, and there was little evidence for their physical existence until fixed target scattering experiments were conducted at the Stanford Linear Accelerator Center in 1968\cite{griff}.
There are 6 of them: up (1968), down (1968), strange (1968), charm (theorized in 1970, discovered 1974), bottom (theorized 1973, discovered 1977), top (theorized 1973, discovered 1995).

Quarks have characteristics such as electric charge, mass, color charge, and spin. The quarks up, charm and top have in common a charge of $\frac{2}{3}$, while the down, strange and bottom quarks have charge of $-\frac{1}{3}$. Among them, the top quark is the most massive of all, while the up and down are the least massive. 
Some of the particles in the SM are not stable, so they decays and form particles with lower masses. Other particles do not decay or their lifetimes are very long. All the particles in the SM are shown in the figure \ref{sm1}. %The final states can be represented with a Feynman diagram according to the SM vertices\cite{mark}.

\begin{center}
  \begin{figure}[ht]
    \centering
    \includegraphics[scale=0.3]{Chapter1/sm1.png}
    \caption[Standard Model to date]{Standard Model to date\cite{smtable}}
    \label{sm1}
  \end{figure}
\end{center}
The last group of the family are the bosons, which have either spin 1 or 0. 
There are 5 bosons: $W^{\pm}$, $Z$ with neutral charge are responsible for weak force; photon, a massless particle, is responsible for the electromagnetic force; gluons which are also massless, is related to the strong force and Higgs boson, recently discovered, is in charge of giving the property of mass to particles via the Higgs mechanism. Bosons let other particles interact, even themselves, which is shown in Figure \ref{sm}. 
The $W$,$Z$ and Higgs bosons, can interact with the most of the particles, while the gluons can only interact with quarks and with itself.
%Those bosons act as mediators in decays and creation of complex
%particles, as mesons and baryons.

%In case of gravity, exist a theoretical particle, called graviton, that is the fundamental particle associated with the gravitational force. In case of its existance, it should have a spin 2 and mass 0.

%There are experiments that detect particles around the world, such as the LHC in Europe, the Fermilab in United States of America, KEK in Japan, etc. More information about the LHC will be described in the next chapters.

	\begin{figure}[ht]
		\centering
		\includegraphics[scale=0.35]{Chapter1/sm.png}
		\caption[Particle interactions in Standard model]{Particle interactions in Standard model\cite{smi}}
		\label{sm}
	\end{figure}

Table \ref{SM table} contains the characteristics of each particle in the SM. The second column shows the masses of the particles, where the most massive particle is the top quark with 173 GeV. There are particles with mass equal to zero, such as the gluons, photons and the neutrinos. 
The third column corresponds to the charge of the particles. For the quarks, there are two different charges: $\frac{2}{3}$ for $u$,$c$ and $t$ and $-\frac{1}{3}$ for $d$,$s$ and $b$. For the leptons, the charge is negative, except for neutrinos, which charge is zero. 
The bosons have a integer charge; $\pm 1$ for $W$ and 0 for $Z$, $\gamma$, $H$ and $g$.	
The next column is the spin, where for fermions (quarks and leptons) is $\frac{1}{2}$ and for bosons is 1\footnote{In general, the fermions have non integer spin and the bosons have integer spin}. The Higgs boson is the only boson with spin 0.
The next one is the lifetime of the particle, where it shows how long the particles exist before decay. Stable lifetime means that the particles do not decay or lifetime almost infinite.
The last column is the distance of particles moving with speeds near light speed. This column illustrates the distance traveled in the predicted lifetime of the particle. And also give us a perspective of how hard the detection of  particles due to their lifetime is.

\begin{table}[htp]
\caption[Table of particles in SM]{SM particle properties: mass, charges, spin, lifetime and distance traveled in a lifetime (assuming speed of $v$=0.998c)\cite{pd}. In the case of $s$,$c$ and $b$ quarks, the lifetime corresponds to the K$^\pm$, D$^{\pm}$ and B$^{\pm}$ meson lifetime.}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|c|p{2.5cm}|p{2.5cm}|}
\hline 
Particle& Mass (MeV/$c^2$) &Charge & spin &Lifetime (s) & Distance (m) \\ 
	\hline 
Up quark ($u$)	& 2.2 & $\frac{2}{3}$ & $\frac{1}{2}$ & Stable & -\\ 
	\hline 
Charm quark ($c$)	& 1280 &$\frac{2}{3}$ & $\frac{1}{2}$ & $ 1.1 \times 10^{-12}$ & 5.21$\times 10^{-3}$ \\ 
	\hline 
Top	quark ($t$)& 173100 & $\frac{2}{3}$ & $\frac{1}{2}$ & 5$\times$10$^{-25}$ &2.37$\times 10^{-15}$  \\ 
	\hline 
Down quark ($d$)	& 4.6 &$-\frac{1}{3}$ & $\frac{1}{2}$ & Stable & - \\ 
	\hline 
Strange quark ($s$)	& 96 &$-\frac{1}{3}$ & $\frac{1}{2}$ &$1.2 \times 10^{-8}$ & 58.7 \\ 
	\hline 
Bottom quark ($b$)	& 4180 &$-\frac{1}{3}$ & $\frac{1}{2}$ &$1.6 \times 10^{-12}$  & 6.16 $\times 10^{-3}$\\ 
	\hline 
$W$ 	& 80379 &$\pm$1 & 1 & 3$\times$ 10$^{-25}$ & 1.4$\times 10^{-15}$\\ 
	\hline 
$Z$ & 91187.6 &0 & 1 & 3$\times$ 10$^{-25}$ &1.4$\times 10^{-15}$ \\ 
\hline
Photon ($\gamma$) & 0 &0 & 1&Stable & - \\ 
\hline
Gluon ($g$)	& 0 &0 & 1&Stable & - \\ 
	\hline 
Higgs ($H$)	& 125.18 &0 & 0& 1.5 $\times$ $10^{-22}$ & 7.39$\times 10^{-13}$ \\ 
	\hline 
Electron ($e$)& 0.511 & -1 &  $\frac{1}{2}$& Stable & - \\ 
	\hline 
Muon ($\mu$)	& 105.7 & -1 & $\frac{1}{2}$ & 2.2$\times$10$^{-6}$ & 10419.85 \\ 
	\hline 
Tau ($\tau$)	& 1776.86 &-1 & $\frac{1}{2}$ & 2.9$\times$10$^{-13}$ & 1.37$\times 10^{-3}$\\ 
	\hline 
$\nu_e$, $\nu_\mu$, $\nu_\tau$&  $<$ 2.0 $\times$ 10$^{-6}$ & 0 & $\frac{1}{2}$ & Stable & -\\
	\hline 
\end{tabular} 
\label{SM table}
\end{table}

\newpage

\section{Higgs Boson}
The Higgs is a boson with spin 0 and chargeless. First theorized in 1964 by a group of theoretical physicist composed by Peter Higgs, Fran\c cois Englert, Robert Brout, Gerald Guralnik, C. Richard Hagen, and Tom Kibble in separate groups. They formulated the Higgs mechanism that explains the generation of mass for the fermions and massive bosons. 
The Higgs boson was discovered recently in year 2012 in the experiments ATLAS and CMS at CERN. The experiment was conducted using proton-proton collisions with center of mass energy of 7 and 8 TeV, where investigators analyzed the data in the $\gamma \gamma$ and $ZZ$ final states\cite{higgsd,Aad_2012}. 
The mass of this boson has been measured and its production and decay
rates are consistent with the SM prediction. The mass of the Higgs boson measured is 125.18 $\pm$ 0.16 GeV\cite{pd}. The Higgs boson lifetime is 1.56 $\times$ $10^{-22}$ s. That is why the Higgs boson was very difficult to detect in particle detectors.

\section{Top quark}
The top quark is the most massive particle with a charge of 2/3 and spin of 1/2. First proposed by M. Kobayashi and T. Maskawa in 1973 to explain observed CP violations in Kaon decay\cite{griff}. In 1995, in the Tevatron at Fermilab in Illinois, the top quark was discovered via proton antiproton collisions with center of mass energy between 1.8 to 1.96 TeV in the CDF and D0 experiments. The top quark production process was via ggF, which generated a pair of $t\bar{t}$\cite{top}.
Since then, the properties of the top quark have been studied at the Tevatron Collider at Fermilab, and by the ATLAS and CMS experiments at the Large Hadron Collider (LHC) at CERN.
However, the top quark often receives special attention in new physics models because its mass requires near unity coupling to the Higgs boson\cite{top}.
As shown in Table \ref{SM table}, the top quark lifetime is so small that they decay before they can hadronise, that is to say, the process to create hadrons from quarks and gluons.

One of the interesting properties of top quark is its mass.
The top quark mass is a free parameter in the SM and was extracted using indirect measurements and relying on SM calculations. The measured value of the top quark mass is of 173 Gev\cite{pd}. The top quark mass is a very important
parameter that puts a constraint in the Higgs boson mass and also takes an important role in electroweak symmetry breaking\cite{top}.

The production of top quark can be generated in different ways. In Tevatron and LHC, the production of $t\bar{t}$ comes mainly from ggF (gluon gluon fusion) via strong force. Even it is possible to create a single top quark in the LHC via electroweak interactions (with a W boson)\cite{th1}. 
The top quark, due to its big mass, always decays into a W boson and a b quark. But since the W boson has various decays, the top quark can generate many possible particles, according to W boson decays. Some possible decay of the top quark is the decay to a lepton such as muon, electron or tau and their neutrinos. %The antitop, antiparticle of top quark, can decay to W$^{-1} \bar{b}$ and its decay is the same as top, but with inverse signal.

\section{Higgs mechanism}
In the SM, mathematically the particles are described in form of fields. The SM for electroweak interactions is described as a gauge theory with a symmetry group SU(2)$\otimes$U(1)
that describes the weak and electromagnetic interactions
due to the exchange of spin 1 gauge fields($W^{\pm}$,$Z$) for the weak part and a photon for the electromagnetic part. The Higgs mechanism give bosons and fermions their mass. The lagrangian for the electroweak part of SM
\begin{align}\label{sml}
\mathcal{L_{\text{EW}}}=\mathcal{L}_\text{gauge}+\mathcal{L}_f +\mathcal{L}_\text{Higgs} + \mathcal{L}_\text{Yukawa}
\end{align}
The first term of the lagrangian refers to gauge bosons 
 \begin{align}\label{smg}
 \mathcal{L}_\text{gauge}=-\frac{1}{4}\left[F^{\mu\nu}F_{\mu\nu}\right]-\frac{1}{4}\left[\sum_{i}G^{i\mu\nu}G^i_{\mu\nu}\right]
 \end{align}
where the F and G are\footnote{$\epsilon^{ijk}$ represents the Levi-Civita tensor, where $\epsilon^{ijk}= \begin{dcases}
	+1 \quad \text{for even permutation} \\
	-1 \quad \text{for odd permutation}\\
	0 \quad \text{i=j, or j=k, or k=i } 
	\end{dcases}$.} 
\begin{align}
F^{\mu \nu}=\partial_\mu B_\nu -\partial_\nu B_\mu \\
G^{i\mu\nu}=\partial_\mu W^i_\nu -\partial_\nu W^i_\mu -g\epsilon^{ijk}W^j_\mu W^k_\nu 
\end{align}
$W^i$ (i=1,2,3) are three SU(2) gauge bosons, B is a U(1) gauge boson and $g$  is a couping constant for $W$ gauge boson. In the end, a superposition of these gauge bosons, after introducing the Higgs mechanism, become the $W^\pm$, $Z$ and $\gamma$ vector boson
\cite{ew1,ew2}. %Z bosons and photons will be formed by a combination of W and B fields.\\

The second term refers to kinetic energy of fermions 
\begin{align}
\mathcal{L}_f=\sum_{m=1}^{3} \left(\bar{q}^m_{L}i \cancel{D}_L q^m_{L}+ 
\bar{l}^m_{L}i \cancel{D}_L l^m_{L} +\bar{u}_{R}^m i \cancel{D}_R u_{R}^m
+\bar{d}_{R}^m i \cancel{D}_R d_{R}^m+\bar{e}_{R}^m i \cancel{D}_R e_{R}^m \right)
\end{align}
where m is the family index, and L(R)
refer to the left (right) handed particles.\footnote{$\bar{\psi}$ denotes the adjoint spinor $\bar{\psi}=\psi^\dagger \gamma^0$.
$\cancel{D}=\gamma^\mu D_\mu$. $\gamma$ are the Dirac matrices}
The gauge covariant derivatives are given by
\begin{align}
{D}_L =\left(\partial_\mu+ig\sigma^iW^i_\mu +ig'Y B_\mu\right)\\
{D}_R =\left(\partial_\mu +ig' Y B_\mu\right)
\end{align}
where $\sigma^i$ are the Pauli matrices and $g'$ is the coupling constant for $B$ gauge boson. %L and R refers to left and right handed fields.
The left handed fields for the first family are defined by the following doublets
\begin{align}
  q_L=\left(\begin{array}{c}
 u_L \\
d_L
\end{array} \right) \qquad  l_L=\left(\begin{array}{c}
 \nu_{eL} \\
e^-_L
\end{array} \right)
\end{align}
and the right handed fields
\begin{align}
  u_R, d_R, e^-_R 
\end{align}
%The charge of particle Q is given by $Q=T^3+Y$ with $Y$ the hypercharge and $T$ the third component of the weak isospin. For left handed doublets $T^3=\pm 1/2$ and $T^3=0$ for right handed singlets and the coefficients.
%g and g' are boson coupling constants\cite{ew2}\\
The hypercharges Y depend of the particle type. For left handed 
\begin{align}
  Y(l_L)=-\frac{1}{2} \quad Y(q_L)=\frac{1}{6} 
 \end{align}
 for right handed 
 \begin{align}
Y(e_R)=-1 \quad Y(u_R)=\frac{2}{3} \quad Y(d_R)=-\frac{1}{3}
\end{align}
The lagrangian for Higgs field
\begin{align}
\mathcal{L}_{\text{Higgs}}=(D_\mu \Phi)^\dagger (D^\mu \Phi)-V(\Phi^\dagger \Phi)
\end{align}
where $D_\mu$ is an operator 
 \begin{align}
D_\mu = \left(\partial_\mu+i\frac{1}{2}\sigma^igW^i_\mu+i\frac{1}{2} g' B_\mu \right) 
 \end{align}
 and the Higgs potential $V(\Phi^\dagger \Phi)$
\begin{align}
 V(\Phi^\dagger \Phi)=\mu^2 \Phi^\dagger \Phi +\frac{1}{2} \lambda (\Phi^\dagger \Phi)^2   
\end{align}
 %is Higgs field which is a complex scalar field that exists in a vacuum and the potential is symmetric under rotations in $\Phi$ space. 
 $\Phi=\left(\begin{array}{c}
 \phi^+ \\
\phi^0
\end{array} \right) $ is a SU(2) doublet of scalar fields. V is symmetrical under rotations in $\Phi$ space
and parameters $\lambda $ and $\mu^2$ are parameters of the potential. The Higgs potential for $\mu^2 <0$ and $\lambda>0$ can be visualized in figure \ref{hat}
%The $\lambda $ term describes self-interactions among the scalar fields. Vacuum stability requires $\lambda $ to be greater than zero. The $\mu$ parameter is a mass parameter for field $\Phi$. $\mu$ and $\lambda$ are Higgs coupling parameters. %If $\mu > 0$, preserve the symmetries of the Lagrangian.
%In the vacuum, the lowest state of energy will be for $\Phi=0$ and the theory is reduced to 
%quantum electrodynamics with a massless photon and a charged scalar
%field $\Phi$ with mass $\mu$.\\
%For $\mu < 0$, the ground state is degenerate and this results in a spontaneous symmetry breaking.

%\begin{figure}[ht!]
%\centering
 %  \begin{tikzpicture}[scale=1.75]
 %   \begin{axis}[
  %     hide axis,
  %    samples=30,
   %    domain=0:360,
   %   y domain=0:1.25
    % ]
    %\addplot3 [surf, shader=flat, draw=blue, fill=white, z buffer=sort] ({sin(x)*y}, {cos(x)*y}, {(y^2-1)^2});
    %\end{axis}
  %\end{tikzpicture}
%  \caption{Higgs potential $V(\Phi^\dagger \Phi)$ for $\mu>0$, also called the mexican hat potential. At the top, for $\mu=0$, there is no mass for the particle.}
%\end{figure}
\begin{figure}[ht]
  \centering
  \begin{tikzpicture}[scale=1.75]
    \begin{axis}[
      hide axis,
      %axis lines=middle,
%      axis on top,
%      axis line style={blue,dashed,thick},
%      ymin=-2,ymax=2,
%      xmin=-2,xmax=2,
%      zmin=-2,zmax=2,
      samples=30,
      domain=0:360,
      y domain=0:1.25,clip=false
    ]
    \addplot3 [surf, shader=flat, draw=black, fill=white, z buffer=sort]
      ({sin(x)*y}, {cos(x)*y}, {(y^2-1)^2});
    \draw[blue,thick,dashed] (axis cs:0,0,0) -- (axis cs:1,0,0)
          node[below,font=\footnotesize]{$\Phi_{\text{IM}}$};
      \draw[blue,thick,-stealth] (axis cs:1,0,0) -- (axis cs:1.3,0,0)
                    node[above,font=\footnotesize]{};
        \draw[blue,thick,dashed] (axis cs:0,0,0) -- (axis cs:0,-1,0)
                    node[left=2mm,font=\footnotesize]{$\Phi_{\text{RE}}$};
        \draw[blue,thick,-stealth] (axis cs:0,-1,0) -- (axis cs:0,-1.5,0)
                    node[right=1mm,font=\footnotesize]{};
        \draw[blue,thick,dashed] (axis cs:0,0,0) -- (axis cs:0,0,1)
                    %node[left=2mm,font=\footnotesize]{$\phi_{\text{RE}}$}
                    ;
        \draw[blue,thick,-stealth] (axis cs:0,0,1) -- (axis cs:0,0,1.3)
                    node[right,font=\footnotesize]{$V(\Phi)$};
        \end{axis}
    \end{tikzpicture}
    \caption{Higgs potential $V(\Phi^\dagger \Phi)$ for $\mu^2<0$ and $\lambda>0$}
    \label{hat}
\end{figure}
Electroweak symmetry breaking refers to the choice of ground state
\begin{align}\label{vmin}
	\Phi_0=\frac{1}{\sqrt{2}}\left(\begin{array}{c}
		0 \\
		v
	\end{array} \right)
	% \Phi =\sqrt{-\frac{\mu^2}{2\lambda}}=\frac{v}{\sqrt{2}}  \qquad \lambda >0
\end{align}
where $v=-\frac{\mu^2}{\sqrt{2}}$ is called the vacuum expectation value.
%Now, since the potential depends on $\Phi^\dagger \Phi$, We can select $\Phi$ in the form 
%\begin{align}\label{doublet}
 %\Phi=\frac{1}{\sqrt{2}}\left(\begin{array}{c}
%0 \\
%v+h
%\end{array} \right) %\equiv v
%\end{align}
%here $\phi^+=0$ and h is a Higgs field with respect to the minimal value v.

%Due to the conservation of electric charge, only a neutral scalar field can acquire a vacuum expectation value.
%With this choice the scalar doublet has U (1)Y charge (hypercharge) $Y_\Phi$= 1
%and the electromagnetic charge is 
%\begin{align}
 %  Q=\frac{1}{2}(\sigma_3+Y)
%\end{align}
%so Q($\phi$) = 0. That is, electromagnetism is unbroken by the scalar VEV. The equation \ref{doublet} breaks the symmetry
%\begin{align}
%SU(2)_L \otimes U(1)_Y \rightarrow U(1)_{em}
%\end{align}
Taking \ref{vmin} and substituting in $\mathcal{L}_\text{Higgs}$, we get 
\begin{align}\label{W}
\begin{split}
%(D_\mu \Phi)^\dagger (D^\mu \Phi) &=\left| \left(\partial_\mu+(ig/2)\sigma^iW^i_\mu+i\frac{1}{2} g' B_\mu \right)\frac{1}{\sqrt{2}}\left(\begin{array}{c}
%0 \\
%v
%\end{array} \right) \right|^2 \notag\\
%&=\frac{v^2}{8}\left| \left(g\sigma^i W^i_\mu + g' B_\mu \right) \left(\begin{array}{c}
%0 \\
%1
%\end{array} \right)    \right|^2 \notag\\
%&=\frac{v^2}{8}\left(\begin{array}{c}
 %  gW^1_\mu-igW^2_\mu \\
 %  -gW^3_\mu+g'B_\mu
%\end{array} \right) \notag\\
(D_\mu \Phi)^\dagger (D^\mu \Phi)& =\frac{v^2}{8} \left(g^2 ((W^1_\mu)^2 +(W^2_\mu)^2 )+( gW^3_\mu -g'B_\mu)^2 \right)
\end{split}
\end{align}

we define the physical vector boson $W^-_\mu$, $W^+_\mu$ and $Z_\mu$ 
\begin{align} 
W^{\pm}_\mu=\frac{1}{\sqrt{2}} (W^1_\mu \mp iW^2_\mu) \label{smo}\\ 
Z_\mu=\frac{1}{\sqrt{g^2+g'^2}}\left(gW^3_\mu -g'B_\mu  \right) \label{zboson}
\end{align}
%From equation \ref{smo}, we can express $W^1_\mu$ and $W^2_\mu$ in terms of $W^{+}_\mu$ and $W^{-}_\mu$
%\begin{align}\label{s2}
%W^{1}_\mu=\frac{W^{+}_\mu+W^{-}_\mu}{\sqrt{2}} \qquad W^{2}_\mu=i\frac{W^{+}_\mu - W^{-}_\mu}{\sqrt{2}}
%\end{align}
%and the mass term
%\begin{align}  
 %  \frac{v^2}{8}g^2 W^\dagger_\mu W^\mu \rightarrow \frac{1}{2} \left( \frac{g v}{2} \right)^2   \notag
%\end{align}
 Now introducing \ref{smo} and \ref{zboson} in \ref{W},
\begin{align}
(D_\mu \Phi)^\dagger (D^\mu \Phi)=\frac{v^2}{8}\left(2g^2 W_\mu^+ W^{\mu-} + (g^2+g'^2)Z_\mu Z^\mu \right) 
\end{align}
from the coefficients, we get the $W$ and $Z$ mass 
\begin{align}
m_W=\frac{gv}{2} \qquad m_Z=\frac{v}{2}\sqrt{g^2+g'^2}
\end{align}
%and the photon that in this case is the field A
%\begin{align}
 %    A_\mu=\frac{1}{\sqrt{g^2+g'^2}}\left(g'W^3_\mu +gB_\mu  \right)
%\end{align}
%with mass 
%\begin{align}
 %     m_A=0
%\end{align}
The last part of the lagrangian is the Yukawa lagrangian\footnote{Here $
	\tilde{\Phi}=i\sigma^2 \Phi^\dagger =\left(\begin{array}{c}
	\phi^{0^\dagger} \\
	-\phi^-
	\end{array} \right)
	% \Phi =\sqrt{-\frac{\mu^2}{2\lambda}}=\frac{v}{\sqrt{2}}  \qquad \lambda >0
	$ is necessary for the correct transformation in the case of up-type quarks. 
	h.c refers to hermitian conjugate terms. } 
\begin{align}
\mathcal{L}_\text{yukawa}=\sum_{m,n}^{3}  \Gamma^u_{mn}\bar{q}_{m\text{,}L} \tilde{\Phi} u_{n\text{,}R}+\Gamma^d_{mn}\bar{q}_{m\text{,}L} \Phi d_{n\text{,}R}+\Gamma^e_{mn}\bar{l}_{m\text{,}L} \Phi e_{n\text{,}R}+h.c
\end{align}
The matrices $\Gamma_{mn}$ describe the so called Yukawa couplings between Higgs doublet $\Phi$ and the fermions. The indices m and n mean sum over the families. 
%It is required to have two representations of Higgs fields
%with $Y=\frac{1}{2}$ and $\frac{1}{2}$ to give masses to quarks and electrons.
By using \ref{vmin} on $\mathcal{L}_\text{yukawa}$, it is obtained for $u$ quark
\begin{align*}
\frac{\Gamma^u_{uu}v}{\sqrt{2}}(\bar{u_L}u_R+\bar{u_R}u_L)
\end{align*}
from which the masses for the fermions are shown to be proportional to the Yukawa couplings
\begin{align*}
m_u=-\frac{\Gamma^u_{uu}v}{\sqrt{2}}
\end{align*}

\pagebreak

\section{Higgs production mechanisms}
During a particle collision, there are various ways that the Higgs boson can be created. The most common is the gluon gluon fusion process ($gg$F) which is the process which involves two gluons mediated by the exchange of a virtual, heavy top.
quarks. Another process is vector boson fusion (VBF), where two fermions interact via a vector boson ($W$ or $Z$ ) and create a Higgs along with other particles. There is also the vector Higgs process (VH), where the creation of particles using vector boson comes from a interaction between fermions and anti fermions\cite{pd}. These processes are shown in the figure \ref{psu}

In the recent years, people have studied the top anti top Higgs ($t\bar{t}H$) process, in order to detect a Higgs boson, via multilepton signals\cite{th1}.
In this process, the creation of a Higgs boson comes from a intection via $t\bar{t}$. In the figure \ref{psu}, the $t\bar{t}H$ starts with 2 gluons which decay to two $t\bar{t}$ pairs and after that, the interaction $t\bar{t}$ generates a Higgs. The last process, subject of study in this thesis, is the top Higgs ($tH$) process. This process is a rare expected process with a very small production rate\cite{pd}. In figure \ref{thw}, another way to produce $tH$, in association with a $W$ is shown.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.5]{Chapter1/pg.png}
\caption[Higgs production mechanism]{Different Higgs production mechanism, from the most likely to least likely \cite{gamma}\cite{timo}}
\label{psu}
\end{figure}

%\begin{figure}[ht]
%\centering
%\includegraphics[scale=0.7]{Chapter1/thw.png}
%\caption[$tHW$ mechanism]{$tHW$ mechanism\cite{th1}}
%\label{thw}
%\end{figure}

In the tH process, a Higgs boson is radiated from a single top quark as shown in Figure \ref{psu}. The $tH$ process has gained interest recently\cite{th1}, given that there are few experiments and its production rate (probability) is small, so the detection of Higgs bosons in this channel opens the path to new discoveries.

In particle physics, cross section ($\sigma$) describes the likelihood of two particles interacting under certain conditions.
 Cross sections are expressed in barns, where 1 barn=$10^{-34}$ cm$^{2}$.
The cross section is important in the evaluation of events for specific processes. In order to get the number of events for a specific process, the reaction rate $N$ is determined by the total cross section $\sigma$ and the luminosity L\footnote{The unit of measurement of instantaneous luminosity is cm$^{-2}$s$^{-1}$. }.
%L is called luminosity, that is the number of particle interactions can be produced in a detector per second.

Therefore, the reaction rate is 
\begin{align} \label{nr}
N_R=\sigma \text{L}
\end{align}
In practice, the luminosity is measured by counting the number of events for a well known process.
\begin{figure}[ht]
	\centering%
	\includegraphics[width=12cm,height=7cm]{Chapter1/7-14xsec.eps}
	\caption[Cross section for different processes generated in a pp collision]{
		Cross section for different processes generated in a pp collision  \cite{dcrosse}}
	\label{csp}
\end{figure}

\begin{table}[ht]
\centering
\caption[Higgs boson production cross sections in pp collisions number of events for  35.9 fb$^{-1}$ ]{	Higgs boson production cross sections in pp collisions for $\sqrt{s}=13$TeV (in pico barn) and number of events for an integrated luminosity of 35.9 fb$^{-1}$ for Run 2  \cite{pd}}
\begin{tabular}{|c|c|c|}
\hline
Production mechanism &
$\sigma$ (picobarns pb) & Number of events \\
\hline
$gg$F & 48.93 & 1756587\\
\hline
VBF & 3.78 & 135702\\
\hline
$WH$ & 1.35 & 48465\\
\hline
$ZH$ &0.88 & 31592\\
\hline
$t\bar{t}$H & 0.50 & 18255\\
\hline
$tH$	& 0.015 & 560.39\\
\hline
\end{tabular}
\label{crt}
\end{table}

 Table \ref{crt} shows the Higgs production cross section for pp collisions and the number of events produced using a integrated luminosity of $35.9 fb^{-1}$, that is the luminosity measured for the Run 2 from LHC on 2016\cite{pd}. Figure \ref{csp} shows the cross sections for each Higgs mechanism process  that comes from pp collisions in relation to colliding energy.
Here we can see that the ggF process has the biggest cross section of all Higgs mechanism and generates the greatest number of events possibles. For the processes tH and $t\bar{t}$H, they have the smallest cross section, so the probability for the processes is low and generates a small number of events. 
\pagebreak

\section{Higgs decay rates}
In particle physics, two important properties are the lifetime and the decay rates of a particle. The lifetime is related to the total decay rate $\Gamma$, that is the probability per unit of time a particle will decay, 
\begin{align}
  dN=-\Gamma N dt
\end{align}
with N the original number of particles before the decay. From here is evident that number of particles left in a decay is 
\begin{align}
N(t)=N_0 e^{-\Gamma t}
\end{align}

%For a particle that decays into several particle, $\Gamma$ is given by
%\begin{align}\label{dm}
%\Gamma=\frac{S}{2\hbar m_n}\int \left| M \right|^2 (2\pi)^4 \delta^4(p_1-p_2-...p_n)\times \prod_{j=2}^n 2\pi \delta (p_j^2-m_j^2 c^2)\theta (p_j^0)\frac{d^4 p_j}{(2\pi)^4}
%\end{align}
%where $m_n$ and $p_n$ are the mass and momentum of the particle n with n=1,2,3,.... S is a statistical factor that avoid identical particles counting. If there are no identical particles, S=1,and for identical particles $S=\frac{1}{s!}$ with s is the number of identical particles. $\mathcal{M}$ is the amplitude which is a momenta function that is calculated acording to the Feynman diagrams.$\theta(p_j^0)$ is a Heaviside function,
%$p_j$ is the outgoing momentum and $p^0_j=E_j/c>0$.\cite{griff}

The number of particles decreases over time exponentially and this can be measured. With this, it is possible determine the mean lifetime\cite{griff}
\begin{align}
\tau=\frac{1}{\Gamma}
\end{align}
In case of the decay rate to a specific process, it is necessary to get the branching fraction of the process. 
The branching ratio for a decay process is the ratio of the number of particles which decay via a specific decay mode with respect to the total number of particles which decay via all decay modes.
\begin{align}
BR_i =\frac{\Gamma_i}{\sum_{i}\Gamma_i}
\end{align}
Where $\Gamma=\sum_i\Gamma_i$ is the total decay width (sum of all partial widths) of the particle.
%Since the dimension of $\Gamma$ is the inverse of time, in our system of natural units, it is measured in inverse seconds, it has the same dimension as mass (or energy)\footnote{Cleaves H.J. (2011) Branching Ratio. In: Gargaud M. et al. (eds) Encyclopedia of Astrobiology. Springer, Berlin, Heidelberg}. 
The lifetime of the Higgs boson is predicted to be 1.56 $\times$ $10^{-22}$ seconds and corresponds to $\Gamma_H$ of about 4 MeV, but has not yet been measured due to the detector resolution\cite{cms-manual}.

\begin{table}[ht] 
\caption[SM Higgs boson branching ratios for $M_H$ =125 GeV]{SM Higgs boson branching ratios for $M_H$ =125 GeV \protect \cite{pd}}
\centering
\begin{tabular}{|c|c|}
\hline
Higgs decay & Branching ratio (BR)\\
\hline
$H \rightarrow$ b$\bar{b}$ &$58.4\%$ \\
\hline
 $H \rightarrow$ $W^+W^-$ &$21.4\%$ \\
\hline
$H \rightarrow$ $\tau^+ \tau^-$ & $6.27\%$\\
\hline
$H \rightarrow$ ZZ &$2.62\%$\\
\hline
$H \rightarrow$ $\gamma\gamma$ &$0.227\%$\\
\hline
$H \rightarrow$ Z$\gamma$ &$0.153\%$\\
\hline
$H \rightarrow$ $\mu^+\mu^-$ &$0.0218\%$\\
\hline
\end{tabular}
\label{higgs1}
\end{table}
The decay $H \rightarrow$ $b\bar{b}$ has a branching ratio a bit more than 50$\%$. The decay $H \rightarrow$ $\mu^+\mu^-$ has a very low branching ratio. This decay is very rare but not impossible to detect in the future. %In this work the majority of the Higgs decays detected are from $WW$, $ZZ$, $\tau \tau$.

\section{$tH$ production mechanism}
The production of $tH$, where a Higgs boson can be radiated
either from the top quark or from the exchanged $W$ boson in the two dominant leading order
diagrams shown in Figure \ref{newth} provides a unique opportunity to study the relative sign of the Higgs top coupling.
In practice, we do not directly measure the Yukawa couplings. Instead, modifier parameters $k_t$ and $k_V$ are introduced in the analysis to represent deviations from the Yukawa coupling values. 
In the SM, the two diagrams interfere negatively and thereby suppress the production cross section.
Any deviation from the SM coupling parameters can lead to a large enhancement of the event rate. 
In this project we consider a modified model with $k_t$=-1, which leads to a constructive interference and a $tH$ cross section more than ten times. 
 %$\kappa_t$ and $\kappa_V$ are the modifier of the coupling parameters discussed in section 1.5, where $\kappa_t$ is for the top and $\kappa_V$ for $W$ boson.

\begin{figure}[ht]
	\centering
	\includegraphics[scale=0.5]{Chapter1/newtHq.png}
	\caption[$tH$ mechanism]{$tH$ mechanism. Higgs radiated from a top quark (left). Higgs radiated from a W boson (right) \protect \cite{bb}} \label{newth}
\end{figure}
\pagebreak
\end{chapter}